\subsubsection{Naïve counting}
If one has taken an elliptic curve $E$ defined by $y^2 = f(x)$ over a finite field $\fp$, a fundamental question to ask is about how many points there are on the curve.
We have already established that there is still a point at infinity on the curve (in Weierstrass form, at $[0,1,0]$), but apart from that, there are a maximum of $p^2$ points.
The naïve approach, of course, is to simply run through all sets of points $(x,y) \in \an[2](\fp)$ and test whether each point satisfies $y^2 = f(x)$, but depending on the size of the field, this may or may not be practical.
Despite this, a basic example of this method of counting is ran through in \cref{hasseweil}.

There are a few approaches possible to take when considering the number of points on an elliptic curve over a finite field, though, and we will cover a number of them.
\subsubsection{The Hasse-Weil estimate}
\label{hasseweil}
The most fundamental theoretical result in finding the order of $E(\fp)$ is the Hasse-Weil theorem, which bounds the size of the group in terms of the prime modulus.
The theorem was initially proved for elliptic curves by Helmut Hasse in \cite{hasse1936a}, \cite{hasse1936b} and \cite{hasse1936c}, and subsequently generalised to curves of higher genus by André Weil in \cite{weil1948}, proved by Pierre Deligne in \cite{deligne1974}.
\input{h-w.tex}
\subsubsection{Schoof's algorithm}
The Hasse-Weil estimate gives an important bound on the number of rational points over an elliptic curve over a finite field, but otherwise does not give the exact number.
For this purpose, in 1985, René Schoof published a paper detailing what is known as \emph{Schoof's algorithm}.
This was the first algorithm of its kind to run in polynomial time, as opposed to exponential time.

We do not go into detail on the algorithm, but mention it for its theoretical significance.
Those who are interested can refer to Schoof's paper detailing the algorithm, \cite{schoof1995}.
\subsubsection{Gauss' theorem}
\input{gauss.tex}
