% This .tex file is the example rsa decryption.
% May need to modify this slightly depending on whether rsa has been introduced at this stage or not.
% Perhaps change the example from year to a key to some other encryption scheme, as would be more realistic.

Now that Lenstra's algorithm has been introduced, we can see a slightly contrived example of a typical use case.
Those unfamiliar with RSA encryption can refer to \cite{menezes1996} for a complete description of the system.
In brief, RSA is a public-key cryptosystem which relies on the difficulty of factoring large integers for its security.
Messages are enciphered with a public-key of the form $(e,N)$, where $e$ is an exponent to raise a number to and $N$ is a (preferably large) semiprime that acts as the modulus for encryption calculations.
Suffice to say, if one can factor the modulus $N$, one can recover the private-key and factor any message enciphered with $(e,N)$, as we demonstrate here.

Say we intercepted a message $M$ along with the public key $P = (e,N)$.
If we were able to factor $N$, we would be able to deduce the decryption key, and thus read this and any future messages.

Let's see how this would work with an example.
Say we knew that
$$ M = 3103229009940552729864,\quad P = (65537, 3160853182090460427047) $$
and we wanted to read $M$.
If we apply Lenstra's algorithm to $N = 3160853182090460427047$, eventually, with $(3,1)$ as our initial point, the algorithm finds that, with $b = 39850$, adding the two points
\begin{align*}
(2191801374392476491053, 2332211434379395076998)\ \text{and}\\
(406058948051076877967, 3156968592727602662096)
\end{align*}
is impossible, since 
$$2191801374392476491053-406058948051076877967 = 1785742426341399613086$$ and $\gcd(1785742426341399613086, 3160853182090460427047) = 3992747141$.
Of course, we don't mind that we can't add the two points, since $3992747141$ is a non-trivial factor of our modulus.
A simple division then gives $N = 3992747141 \times 791648724667$.
Because we know this is an RSA modulus, we know that this is the complete factorisation of $N$, but for completeness, a simple brute force check quickly shows that $p = 791648724667$ and $q = 3992747141$ are both prime.

Now that we know this, we calculate $(p-1) \times (q-1) = 3160853181294818955240$ and then find the multiplicative inverse of $e = 65537$ modulo $3160853181294818955240$ with the euclidean algorithm.
This turns out to be $146040609624497792033$, and we finally decrypt our message by taking
$$ M^{146040609624497792033} \mod 3160853182090460427047 =2016 $$
so we see that whatever nefarious organisation we are spying on is communicating the current year.
